---
title: "Visualize spatial rainfall of each province in Thailand (December, 2025) "
caption: using plotly and R Markdown presentation
author: "Naddao Thongsak"
date: "`r Sys.Date()`"
output:
  ioslides_presentation: default
  slidy_presentation: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results=TRUE)
```

# 1. Data sources

## Spatial rainfall data of all provinces of Thailand 

**Source:** Thailand spatial rainfall data is downloaded from website of Open
Goverment data of Thailand (<https://data.go.th/dataset/spatial-rain>).

**Data description:** Data is described as spatial rainfall of each province in Thailand
summary for each month since 2018 to 2026. Spatial rainfall is an
average of rainfall of each interested area (such as province). Spatial rainfall data combine information of
rainfall data from multiple points station(station that measure amount of
rainfall) then use algorithm to interpolate data into data that represent rainfall amount for that area.

# 2. Understanding our data

## 2.1 Read spatial rainfall data of Thailand from .csv file.

```{r, echo = TRUE}
# load spatial rainfall data to memory
url <- "spatial-rain-hii.csv"
spt_rainfall_th <- read.csv(url)
```

View first 6 rows of data.

```{r, echo = TRUE}
head(spt_rainfall_th)
```

## 2.2  Overview spatial rainfall data of Thailand (2018-2026)

**Data dimension:** From spatial rainfall data of Thailand 2018-2026,
data contains 7546 rows and 7 columns.

```{r, echo = TRUE}
dim(spt_rainfall_th)
```

**Data period**

```{r, echo = TRUE}
unique(spt_rainfall_th$YEAR)
```

## 2.3 Understanding data in each columns (#1)

**data in each column** are the following:

-   YEAR: year that collect data.

-   MONTH: month that collect data

-   PROV_ID: ID of each province in Thailand

-   PROV_T: province name in Thai

## 2.3 Understanding data in each columns (#2)

-   MinRain: Minimum spatial rainfall data of specify province in that month and year.

-   MaxRain: Maximum spatial rainfall data of specify province in that month and year.

-   AvgRain : Average spatial rainfall data of specify province in that month and year.

## 2.4 Understanding data in each row

**Data in each row** is a spatial rainfall data of each province in Thailand of specify month, year. There are 3 measurement rainfall data which are:-

-   AvgRain: Avergage Rainfall(mm)

-   MinRain: Maximum rainfall(mm)

-   MaxRain: Minimum rainfall(mm)

```{r, echo = TRUE, results=TRUE}
head(spt_rainfall_th, n=1)
```

# 3. Prepare data of spatial rainfall for all province of Thailand in December, 2025

## 3.1 Filter only interested period, December, 2025.

Filter spatial rainfall data of all province in Thailand using only data of December, 2025.

In this example, we will use spatial rainfall data on December, 2025 for all provinces in Thailand. We use R library "dplyr" to help filter data.

```{r, echo = TRUE, results=TRUE}
library(dplyr)
```

## 

```{r, echo = TRUE, results=TRUE}
## Filter Thailand rainfall of each province in Thailand only in 2025
spt_rainfall_2025 <- spt_rainfall_th %>%
                          filter(spt_rainfall_th$YEAR==2025)

## Filter Thailand rainfall of each province in Thailand 
#only in December, 2025
spt_rainfall_2025_12 <- spt_rainfall_2025 %>%
                            filter(spt_rainfall_2025$MONTH==12)
```

## 3.2 Sorting data using column Average Rainfall(AvgRain)

For ploting the graph, we want to see data of Average Rainfall sorting from province which has maximum average rainfall down to province which
has minimum average rainfall, then we will sort our data using data of Average rain(AvgRain).

```{r, echo=TRUE}
spt_rainfall_2025_12 <- spt_rainfall_2025_12 %>%
                            arrange(desc(AvgRain))
spt_rainfall_2025_12$PROV_T <-  factor(
                      spt_rainfall_2025_12$PROV_T,
                          levels = spt_rainfall_2025_12$PROV_T)
head(spt_rainfall_2025_12, n=3)
```

# 4. Plot data using scatter plot for all province of Thailand in December, 2025
##  
In this section, we plot data of Average Rainfall, Minimum Rainfall and Maximum rainfall of
each province in Thailand on December, 2025, using plotly

## 4.1 For visualization graph, we use R library "plotly".

```{R, echo=TRUE, results=FALSE}
library(plotly)
```

## 4.2 Average rainfall, Minimum rainfall, Maximum rainfall as data for each line graph

There is 3 columns of data which we use in this plot:-

-   Average Rainfall(AvgRain), mm.
-   Minimum Rainfall(Minimum Rainfall) mm.
-   Maximum Rainfall(Maximum Rainfall) mm.

## 4.3 Using scatter plot(lines and marker)

We visualize plotly graph using scatter plot(lines and markers).

```{R, echo=TRUE, results=FALSE}
fig1 <- plot_ly(spt_rainfall_2025_12,
                x = ~PROV_T,
                y = ~AvgRain,
                type = "scatter",
                mode="markers+lines",
                name="Average Rainfall (mm.)") %>%
          add_trace(y=~MinRain,
                    name="Minimum Rainfall(mm.)",
                    mode="markers+lines" ) %>%
          add_trace(y=~MaxRain,
                    name="Maximum Rainfall(mm.)",
                    mode="markers+lines" )  %>%
          layout(title = "Spatial rainfall from each province in Thailand (December 2025)",
                 xaxis = list(title = "Province Name(in THAI)"),
                 yaxis = list (title = "Rainfall (mm)") )
```

## 4.4 Visualize graph

```{R, echo=TRUE, results=TRUE}
fig1
```

# 5. Prepare data of spatial rainfall for Phuket province in Thailand in year 2025.

##   
In this section, We will prepare data of spatial rainfall for Phuket province of Thailand in year 2025.

## 5.1 Filter data for only Phuket("ภูเก็ต" in Thai) province.

```{r, echo = TRUE}
## Thailand rainfall of each province in Thailand(2025)
spt_rainfall_phuket_2025 <- spt_rainfall_2025 %>%
                               filter(spt_rainfall_2025$PROV_T=="ภูเก็ต")
```

## 5.2 Sorting data from the first month to last month of the year.

```{r, echo=TRUE}
spt_rainfall_phuket_2025 <- spt_rainfall_phuket_2025 %>%
                              arrange(MONTH)

head(spt_rainfall_phuket_2025, n=6)
```

# 6. Plot data using scatter plot for Phuket province in Thailand in year 2025.
## 6.1 Create ploting data using R library "plotly" to visualize data
spatial rainfall of Phuket province of Thailand in year 2025.
```{R, echo=TRUE, results=FALSE}
fig2 <- plot_ly(spt_rainfall_phuket_2025,
                x = ~MONTH,
                y = ~AvgRain,
                type = "scatter",
                mode="markers+lines",
                name="Average Rainfall(mm.)") %>%
  layout(title = "Spatial rainfall of Phuket, Thailand (all month in 2025)",
                 xaxis = list(title = "Month"),
                 yaxis = list (title = "Rainfall (mm)") ) %>%
          add_trace(y=~MinRain,
                    name="Minimum Rainfall(mm.)",
                    mode="markers+lines" ) %>%
          add_trace(y=~MaxRain,
                    name="Maximum Rainfall(mm.)",
                    mode="markers+lines" ) 
```

## 6.2 Visualization plot

```{R, echo=TRUE, results=TRUE}
fig2
```

# 7. Plot spatial rainfall data of all provinces Thailand December, 2025 and visualize in choropleth map.

##  
In this section, we use choropleth map of library plotly to visualize data as map to show
amount of Average rainfall(AvgRain) for all provinces in Thailand 

## 7.1 Using R library "rjson" to read GeoJSON data.

```{R, echo=TRUE, results=FALSE}
library(rjson)
```

## 7.2 Read GeoJSON data into memory.

```{R, echo=TRUE, results=FALSE}
url <- "simplified_provinces_geojson.geojson"
geojson <- rjson::fromJSON(file=url)
```

## 7.3 Configure the appearance of the choropleth map

```{R, echo=TRUE}
g <- list(
  fitbounds = "locations",
  visible = TRUE
)
```

## 7.4 Create choropleth map
```{R, echo=TRUE, results=FALSE}
fig_avg_rain <- plot_ly()  %>% 
                  add_trace(
                    type="choropleth",
                    geojson=geojson,
                    locations=spt_rainfall_2025_12$PROV_T,
                    z=spt_rainfall_2025_12$AvgRain,
                    colorscale="Viridis",
                    featureidkey="properties.ADM1_TH" ) %>% 
                  layout( geo = g ) %>% 
                  colorbar(title = "Average Rainfall (mm)" ) %>% 
                  layout(
                    title = "Amount of Average Rainfall(AvgRain) of provinces in Thailand in December, 2025" )
```

## 7.5 Visualize the choropleth map
Visualization describe spatial rainfal(average rain, AvgRain) of all province in Thailand in December, 2025
```{R, echo=FALSE, results=TRUE}
fig_avg_rain
```
